Donor Preferences in Canadian Political Financing

### 2024-12-04 paper/paper.qmd
update discussion heading

## The Influence of Party Size on Donation Amounts {#sec-first-point}

@@

The landscape in Canada for political financing is growing each year.
Results from studies demonstrate that individual contributors may prefer
to support challengers for several reasons. An optimistic but naive
explanation may be that individuals hope to foster competitive
elections, given that donations depend not on political leaning but on
who is currently in power. However, a more realistic explanation for
this phenomenon is that some individuals want to make policy changes. It
is no secret that politicians, particularly those out of power, often
solicit funding and political support in exchange for advancing the
interests of their stakeholders @mollona2022ethics}. These two dynamics
may be working in tandem to create an effect where parties eager to be
elected rely more on funding from external sources, which can switch
each election cycle.

Moreover, these trends of political donations fluctuating based on the
party in power are not unique to Canada. The study "Financing UK
Democracy: A Stocktake of 20 Years of Political Donations" highlights a
surge in private political donations, more specifically from
superdonors, widening the resource gap between large parties
@draca2022financing}. Similar to Canada, the UK also sees large
donations from a wealthy few influencing their elections each year. The
parallel between Canada and the UK underscores a broader democratic
issue of financial power being used to influence political outcomes.
Though this concept is not novel, this study contributes to the already
growing spotlight on the persistent lack of proper reform in political
systems.

>>>>

## Financial influence shapes political landscapes

@@

The landscape in Canada for political financing is growing each year.
Results from studies demonstrate that individual contributors may prefer
to support challengers for several reasons. An optimistic but naive
explanation may be that individuals hope to foster competitive
elections, given that donations depend not on political leaning but on
who is currently in power. However, a more realistic explanation for
this phenomenon is that some individuals want to make policy changes. It
is no secret that politicians, particularly those out of power, often
solicit funding and political support in exchange for advancing the
interests of their stakeholders @mollona2022ethics. These two dynamics
may be working in tandem to create an effect where parties eager to be
elected rely more on funding from external sources, which can switch
each election cycle.

Moreover, these trends of political donations fluctuating based on the
party in power are not unique to Canada. The study "Financing UK
Democracy: A Stocktake of 20 Years of Political Donations" highlights a
surge in private political donations, more specifically from
superdonors, widening the resource gap between large parties
@draca2022financing. Similar to Canada, the UK also sees large
donations from a wealthy few influencing their elections each year. The
parallel between Canada and the UK underscores a broader democratic
issue of financial power being used to influence political outcomes.
Though this concept is not novel, this study contributes to the already
growing spotlight on the persistent lack of proper reform in political
systems.
